Segregation in Vienna's primary schools?

### 2021-01-13 _post/2020-03-02-primary-schools-in-vienna/primary-schools-in-vienna.Rmd
adding revised school plot; crosstalk visible now

---
title: "Segregation in Vienna's primary schools?"
description: |
  A short description of the post.
date: 03-02-2020

@@

<details closed> 
<summary>Click here for code on how to geolocate schools' adresses </summary>
```{r, eval=F}

# apply function google_geocode to each address
# use map (not map_chr) since you'll get a bunch of geographic details
df_primary_schools_vienna <- df_primary_schools_vienna %>%
  mutate(coord = map(
    school_adress,
    ~ google_geocode(.x, key = your_google_api_key)
  ))

@@

```{r, eval=T}
# add details of schools to dataframe containing all combinations of schools (nested);
# first for 1st school in pair, then for 2nd school in pair;
# extract values of interest for easier handling from nested dataframe;
df_main <- df_difference %>%
  mutate_at(vars("school_1", "school_2"), as.numeric) %>%
  nest_join(., df_primary_schools_vienna,
    by = c("school_1" = "properties_skz")
  ) %>%
  rename(school_1_details = df_primary_schools_vienna) %>%
  mutate(school_1_lat = map_dbl(school_1_details, purrr::pluck, "lat")) %>%
  mutate(school_1_long = map_dbl(school_1_details, purrr::pluck, "long")) %>%
  # nest(school_1_coords=c(school_1_lat, school_1_long)) %>%
  nest_join(., df_primary_schools_vienna,
    by = c("school_2" = "properties_skz")
  ) %>%
  rename(school_2_details = df_primary_schools_vienna) %>%
  mutate(school_2_lat = map_dbl(school_2_details, purrr::pluck, "lat")) %>%
  mutate(school_2_long = map_dbl(school_2_details, purrr::pluck, "long")) %>%
  mutate(address1 = map(school_1_details, ~ .x %>% mutate(school_1_address = paste(properties_bezeichnung,
    properties_str,
    properties_plz,
    sep = ", "
  ))) %>%
    map_chr(., purrr::pluck, "school_1_address")) %>%
  mutate(address2 = map(school_2_details, ~ .x %>% mutate(school_2_address = paste(properties_bezeichnung,
    properties_str,
    properties_plz,
    sep = ", "
  ))) %>%
    map_chr(., pluck, "school_2_address")) %>%
  mutate(school_1_erhalter = map(school_1_details, purrr::pluck, "properties_erhalter")) %>%
  mutate(school_2_erhalter = map(school_2_details, purrr::pluck, "properties_erhalter")) %>%
  mutate(school_erhalter_indicator = map2_chr(
    school_1_erhalter, school_2_erhalter,
    function(x, y) paste(pmin(x, y), pmax(x, y), sep = "-")
  )) %>%
  mutate(school_1_erhalter = map(school_1_details, extract, "properties_erhalter") %>% unlist()) %>%
  mutate(school_2_erhalter = map(school_2_details, extract, "properties_erhalter") %>% unlist()) %>%
  mutate(erhalter_type = paste(school_1_erhalter, school_2_erhalter, sep = " - ")) %>%
  left_join(., df_share %>%
    filter(school_type_secondary == "matura") %>%
    select(school_1_n_rel = n_rel, properties_skz_vj),
  by = c("school_1" = "properties_skz_vj")
  ) %>%
  left_join(., df_share %>%
    filter(school_type_secondary == "matura") %>%
    select(school_2_n_rel = n_rel, properties_skz_vj),
  by = c("school_2" = "properties_skz_vj")
  )
```

@@

# cacluate the factor by which one school's AHS differeds from another school's AHS ratio.
# the pmin/pmax function ensures that the school with the lower AHS ratio is
# is the denominator; makes comparison between school pairs easier;
df_main <- df_main %>%
  mutate(diff_factor = map2_dbl(
    school_1_n_rel, school_2_n_rel,
    function(x, y) pmax(x, y) / pmin(x, y)
  ))
```

@@

```{r, eval=T}

# take only those schools located within 500 m vicincity;
# and only public school dyads

@@

cross_map2 <- plotly::plot_mapbox(
  data = shared_df_main,
  hoverinfo = "text",
  text = ~ paste(
    "abs diff AHS ratio (% points):", round((diff_abs * 100), 2),
    "</br>AHS factor:", round(diff_factor, 2),
    "</br>distance (m):", round(distance, 0),
    "</br>", tooltip
  )
) %>%
  plotly::add_markers(
    y = ~school_1_lat,
    x = ~school_1_long,
    color = "#FF0000" # ,
    # text = ~address1
  ) %>%
  plotly::add_markers(
    y = ~school_2_lat,
    x = ~school_2_long,
    color = "#FF0000" # ,
    # text = ~address2
  ) %>%
  layout(
    mapbox = list(
      style = "mapbox://styles/zoowalk/ck6kz867d0jdi1io2lxyp9008",
      zoom = 10,
      center = list(
        lat = 48.21,
        lon = 16.3720798
      )
    ),
    showlegend = FALSE,
    paper_bgcolor = bgr_color
  )

# combine graphs

# cross_plot <- crosstalk::bscols(
# plot_distance_diff,
# cross_map2)

library(htmltools)
cross_plot <- browsable(tagList(plot_distance_diff, cross_map2))

# see here for how to remove bootstrap css which otherwise messes up you blogdown's css
# https://github.com/rstudio/crosstalk/issues/26

@@

cross_plot_clean <- unstrap(cross_plot)
```
</details>

```{r echo=FALSE}

@@

What the data does not tell us is how this discrepancy actually came about. By the best of my imagination, I find it difficult to believe that there is some pro-active, intentional policy going on from the side of the authorities. What seems more likely to me - and I can only speculate - is that we are witnessing the result of a self-perpetuating dynamic. With more German native speakers attending the school in one school, AHS ratios are higher. Consequenlty, German speaking parents send their kids again to the school with the higher AHS ratio and not to the other, similarly close school. Obviously, this still leaves a lot of other questions unaswered (e.g. (why) do non-German speaking parents exhibit the same enrollment behavior as German speaking parents?). However, what seems clear to me though is that this dynamic would warrant some intervention to lift up the schools with the lower AHS ratios.

>>>>

---
title: "Segregation in Vienna's primary schools?"
description: |
  Drawing on public data from Vienna's school administration I look at how many percentages of pupils in each primary schoos subsequently leave for schools offering eventually a higher education entrance qualification (Matura). The main interest is in contrasting these ratios between closely (geo-)located schools.

date: 03-02-2020

theme: theme.css

@@

<details closed> 
<summary>Click here for code on how to geolocate schools' addresses </summary>
```{r, eval=F}

# apply function google_geocode to each address
# use map (not map_chr) since you'll get a bunch of geographic details

df_primary_schools_vienna <- df_primary_schools_vienna %>%
  mutate(coord = map(
    school_adress,
    ~ google_geocode(.x, key = your_google_api_key)
  ))

@@

```{r, eval=T}
# add details of schools to dataframe containing all combinations of schools (nested);
# first for 1st school in pair, then for 2nd school in pair;
# extract values of interest for easier handling from nested dataframe;
df_main <- df_difference %>%
  mutate_at(vars("school_1", "school_2"), as.numeric) %>%
  nest_join(., df_primary_schools_vienna,
    by = c("school_1" = "properties_skz")
  ) %>%
  rename(school_1_details = df_primary_schools_vienna) %>%
  mutate(school_1_lat = map_dbl(school_1_details, purrr::pluck, "lat")) %>%
  mutate(school_1_long = map_dbl(school_1_details, purrr::pluck, "long")) %>%
  # nest(school_1_coords=c(school_1_lat, school_1_long)) %>%
  nest_join(., df_primary_schools_vienna,
    by = c("school_2" = "properties_skz")
  ) %>%
  rename(school_2_details = df_primary_schools_vienna) %>%
  mutate(school_2_lat = map_dbl(school_2_details, purrr::pluck, "lat")) %>%
  mutate(school_2_long = map_dbl(school_2_details, purrr::pluck, "long")) %>%
  mutate(address1 = map(school_1_details, ~ .x %>% mutate(school_1_address = paste(properties_bezeichnung,
    properties_str,
    properties_plz,
    sep = ", "
  ))) %>%
    map_chr(., purrr::pluck, "school_1_address")) %>%
  mutate(address2 = map(school_2_details, ~ .x %>% mutate(school_2_address = paste(properties_bezeichnung,
    properties_str,
    properties_plz,
    sep = ", "
  ))) %>%
    map_chr(., pluck, "school_2_address")) %>%
  mutate(school_1_erhalter = map(school_1_details, purrr::pluck, "properties_erhalter")) %>%
  mutate(school_2_erhalter = map(school_2_details, purrr::pluck, "properties_erhalter")) %>%
  mutate(school_erhalter_indicator = map2_chr(
    school_1_erhalter, school_2_erhalter,
    function(x, y) paste(pmin(x, y), pmax(x, y), sep = "-")
  )) %>%
  mutate(school_1_erhalter = map(school_1_details, extract, "properties_erhalter") %>% unlist()) %>%
  mutate(school_2_erhalter = map(school_2_details, extract, "properties_erhalter") %>% unlist()) %>%
  mutate(erhalter_type = paste(school_1_erhalter, school_2_erhalter, sep = " - ")) %>%
  left_join(., df_share %>%
    filter(school_type_secondary == "matura") %>%
    select(school_1_n_rel = n_rel, properties_skz_vj),
  by = c("school_1" = "properties_skz_vj")
  ) %>%
  left_join(., df_share %>%
    filter(school_type_secondary == "matura") %>%
    select(school_2_n_rel = n_rel, properties_skz_vj),
  by = c("school_2" = "properties_skz_vj")
  )

# readr::write_rds(df_main, file=here::here("_blog_data", "Vienna_primary_schools","df_main2.rds"))

@@

# cacluate the factor by which one school's AHS differeds from another school's AHS ratio.
# the pmin/pmax function ensures that the school with the lower AHS ratio is
# is the denominator; makes comparison between school pairs easier;

df_main <- df_main %>%
  mutate(diff_factor = map2_dbl(
    school_1_n_rel, school_2_n_rel,
    function(x, y) pmax(x, y) / pmin(x, y)
  ))
```

@@

```{r, eval=T}
# take only those schools located within 500 m vicincity;
# and only public school dyads

@@

cross_map2 <- plotly::plot_mapbox(
  data = shared_df_main,
  hoverinfo = "text",
  text = ~ paste(
    "abs diff AHS ratio (% points):", round((diff_abs * 100), 2),
    "</br>AHS factor:", round(diff_factor, 2),
    "</br>distance (m):", round(distance, 0),
    "</br>", tooltip
  )
) %>%
  plotly::add_markers(
    y = ~school_1_lat,
    x = ~school_1_long,
    color = "#FF0000" #,
    # text = ~address1
  ) %>%
  plotly::add_markers(
    y = ~school_2_lat,
    x = ~school_2_long,
    color = "#FF0000" # ,
    # text = ~address2
  ) %>%
  layout(
    mapbox = list(
      style = "mapbox://styles/zoowalk/ckjrkv7065tf519oax4nadm2k",
      margin=list(l=0,r=0,t=0,b=0,p=0),
      zoom = 10,
      center = list(
        lat = 48.21,
        lon = 16.3720798
      )
    ),
    showlegend = FALSE,
    paper_bgcolor = bgr_color
  ) %>% 
    config(mapboxAccessToken = Sys.getenv("MAPBOX_TOKEN"))
```
</details>

```{r, eval=T}
# combine graphs

cross_plot <- crosstalk::bscols(
plot_distance_diff,
cross_map2)
#

cross_plot <- crosstalk::bscols(
plot_distance_diff,
cross_map2)
```{r eval=FALSE, include=FALSE}
# cross_plot
```

cross_plot <- browsable(tagList(cross_plot))
cross_plot

```{r, eval=T}
library(htmltools)
# cross_plot <- browsable(tagList(plot_distance_diff, cross_map2))
cross_plot <- browsable(tagList(cross_plot))
# cross_plot
```

```{r, eval=T}
# see here for how to remove bootstrap css which otherwise messes up you blogdown's css
# https://github.com/rstudio/crosstalk/issues/26

@@

cross_plot_clean <- unstrap(cross_plot)
```

cross_plot_clean
```{r echo=FALSE, layout="l-page", fig.height=30}

@@

What the data does not tell us is how this discrepancy actually came about. By the best of my imagination, I find it difficult to believe that there is some pro-active, intentional policy going on from the side of the authorities. What seems more likely to me - and I can only speculate - is that we are witnessing the result of a self-perpetuating dynamic. With more German native speakers attending the school in one school, AHS ratios are higher. Consequently, German speaking parents send their kids again to the school with the higher AHS ratio and not to the other, similarly close school. Obviously, this still leaves a lot of other questions unanswered (e.g. (why) do non-German speaking parents exhibit the same enrollment behavior as German speaking parents?). However, what seems clear to me though is that this dynamic would warrant some intervention to lift up the schools with the lower AHS ratios.
